TRIK Studio: Technical Introduction

### 2017-02-10 2017-trikStudioEnglish/trikStudioEnglish.tex
More polishing

Almost all aforementioned visual languages are based on control-flow computational model. Such a model is easier to comprehend, so it suits the purposes of education well. Nevertheless this model is not the best for programming robots since they are highly reactive by their nature: a program controlling a robot is basically a transformation of sensor signals into motor impulses. Reactive models are best expressed in data-flow languages~\cite{johnston2004advances}, where a program consists of a number of "`black boxes"', connected with data channels. Each such "`box"' (we will call it \emph{block}) has a fixed number of inputs and outputs, and its job is to transform input data into output data (we will call \emph{tokens} data transferring through channels). For instance, each robot's sensor is represented by a single block, that simply sends tokens with sensor data to its output. Many researchers  note usability of visual data-flow languages comparing to control-flow ones~\cite{johnston2004advances}, in particular because of clear visualization of data flows.

@@

\verb|LuaTokenTypes::semicolon| is a token corresponding to ';', operator '-' creates a simple parser that can parse only semicolons and excludes semicolons from the AST, "`explist"` is a reference parser object, like "`stat"', defined elsewhere, '\&' combines two parsers into a parser which accepts concatenation of their corresponding strings, '|' combines two parsers into a parser that accepts alternative, '\verb|~|' creates an optional parser from its argument, which shall also be a parser. There are also operators for adding semantic actions to productions and assigning parser a name for debug purposes. When all parsers are combined in such a way, it is enough to call 'parse()' method of a resulting parser object, giving it a token stream. The parser combinator library was used for another language in QReal~\cite{tikhonova2015generation}, so it is general enough to support not only Lua.

>>>>

Almost all aforementioned visual languages are based on control-flow computational model. Such a model is easier to comprehend, so it suits the purposes of education well. Nevertheless this model is not the best for programming robots since they are highly reactive by their nature: a program controlling a robot is basically a transformation of sensor signals into motor impulses. Reactive models are best expressed in data-flow languages~\cite{johnston2004advances}, where a program consists of a number of "`black boxes"', connected with data channels. Each such "`box"' (we will call it \emph{block}) has a fixed number of inputs and outputs, and its job is to transform input data into output data (we will call \emph{tokens} data transferring through channels). For instance, each robot's sensor is represented by a single block, that simply sends tokens with sensor data to its output. Many researchers note usability of visual data-flow languages comparing to a textual ones~\cite{johnston2004advances}, in particular because of clear visualization of data flows.

@@

\verb|LuaTokenTypes::semicolon| is a token corresponding to ';', operator '-' creates a simple parser that can parse only semicolons and excludes semicolons from the AST, "`explist"' is a reference parser object, like "`stat"', defined elsewhere, '\&' combines two parsers into a parser which accepts concatenation of their corresponding strings, '|' combines two parsers into a parser that accepts alternative, '\verb|~|' creates an optional parser from its argument, which shall also be a parser. There are also operators for adding semantic actions to productions and assigning parser a name for debug purposes. When all parsers are combined in such a way, it is enough to call 'parse()' method of a resulting parser object, giving it a token stream. The parser combinator library was used for another language in QReal~\cite{tikhonova2015generation}, so it is general enough to support not only Lua.
